---
title: "RepData"
author: "Csaba Hurton"
date: "23/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

## Loading and preprocessing the data

###### 1. Load the data
```{r loading}
data <- read.table(unz("activity.zip", "activity.csv"), header=TRUE, quote="\"", sep=",")

```

###### 2. Process/transform the data (if necessary) into a format suitable for your analysis
- Looking into the data, I saw rows where steps is "NA". I removed them with the help of the "tidyr" package
- Also, date is type "chr", I converted it to a date format
```{r noNA}
head(data)
library(tidyr)
data <- drop_na(data)

data$date <- as.Date(data$date)

```
## What is mean total number of steps taken per day?
###### 1. Calculate the total number of steps taken per day
```{r}
steps_per_number <- aggregate(data["steps"], by=data["date"], sum)
steps_per_number
```

###### 2. Make a histogram of the total number of steps taken each day
```{r}
hist(steps_per_number$steps,
     main="Histogram: total number of steps taken each day", 
     xlab="steps", 
     breaks=20)
```

###### 3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mn <- round(mean(steps_per_number$steps), digits = 2)
mdn <- median(steps_per_number$steps)

mean_median_df <- data.frame()
mean_df <- aggregate(data["steps"], by=data["date"], mean)
names(mean_df)[names(mean_df) == "steps"] <- "mean"

median_df <- aggregate(data["steps"], by=data["date"], median)
mean_df['median'] <- median_df['steps']
mean_df

```

## What is the average daily activity pattern?

###### 1. Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
avg_by_interval <- aggregate(data["steps"], by=data["interval"], mean)

plot(avg_by_interval$interval, avg_by_interval$steps, 
     type = "l", 
     main="avg # of steps by 5 min interval", 
     xlab="interval",
     ylab="avg steps")
```
###### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
mx = max(avg_by_interval$steps)
temp <- avg_by_interval[avg_by_interval[, "steps"] == max(avg_by_interval$steps),]
ntrvl <- as.numeric(temp["interval"])
ntrvl
```


The **`r ntrvl`**th 5-minute interval contains the maximum number of steps.

#### Imputing missing values

###### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
data2 <- read.table(unz("activity.zip", "activity.csv"), header=TRUE, quote="\"", sep=",")
nas <- sum(is.na(data2))

```
###### 2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. 
the mean for that 5-minute interval, etc.
###### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
data2[is.na(data2[,"steps"]),"steps"] <- as.integer(avg_by_interval$step+0.5)
head(data2)
```
###### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
steps_per_number2 <- aggregate(data2["steps"], by=data2["date"], sum)

hist(steps_per_number2$steps,
     main="Histogram: total number of steps taken each day", 
     xlab="steps", 
     breaks=20)
```
Modifying the dataset kept the tendencies but increased the daily steps.

## Are there differences in activity patterns between weekdays and weekends?

###### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
As the task does not specify, I am going to use the original "data" dataframe that omits the NA values. "data2" made some generalization by replacing NAs with mean values - weekday data is put to weekends and weekend average is put to weekdays.
```{r}
data["weekday"] <- weekdays(as.Date(data$date))

data[data[,"weekday"] %in% c("Saturday", "Sunday"), "weekday"] <- "weekend"
data[data[,"weekday"] %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), "weekday"] <- "weekday"

avg_by_interval <- aggregate(data["steps"], by=data[c("interval", "weekday")], mean)

```
###### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
library(lattice)
xyplot(steps ~ interval | weekday, data = avg_by_interval, layout = c(1,2), type="l", xlab="Interval", ylab="Number of steps")
```

